---
title: "CDEC Overview"
author: "[Skyler Lewis](mailto:slewis@flowwest.com)"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    theme: flatly
  github_document:
    default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(sf)
```

## CDEC Summary 

[1 sentence description of data type (modeled or gage)...and source and purpose.]

* **Source:** California Department of Water Resources (DWR), aggregating hydrologic datasets from many sources including counties, water districts, CDFW, NPS, USFS, USGS, USACE, and DWR itself 
* **Accessibility:** Public, open and accessible online
* **Coverage:** State of California, where data is available
* **Temporal Coverage:** [max temporal range (min date - max date)] - describe variation across gages if relevant 
* **Spatial Coverage:** [x out of x watersheds] 
* **Maintenance:** Is it maintained? And how often? By who (if different than source)
* **Contact: ** Data / Model Contact (think about if worth expanding to a larger list of people / orgs) - think about having table section below with projects/contacts 
* **Utilized By:** What process is it used in: (list processes that use this data) 

## Data Access

### Parameters

#### Stations

Station IDs follow a 3-digit format

Station Lookup: `https://cdec.water.ca.gov/dynamicapp/staMeta?station_id={XXX}`

Station locator map: [https://cdec.water.ca.gov/webgis/?appid=cdecstation](https://cdec.water.ca.gov/webgis/?appid=cdecstation)

Stations can be searched by river basin

#### Sensors

Refer to the station ID for which sensors are available at a particular station.

Following are the sensor codes that report some variation of flow in cfs. There are many other codes not listed here including stage, temperature, turbidity, etc.

* **`20`** = **FLOW** = FLOW, RIVER DISCHARGE, CFS
* **`41`** = **MFLOW** = FLOW, MEAN DAILY, CFS
* **`165`** = **FLOW.XX** = FLOW, RIVER DISCHARGE PRECISE, CFS (rarely used)
* **`76`** = **INFLOW** = RESERVOIR INFLOW, CFS
* **`23`** = **OUTFLOW** = RESERVOIR OUTFLOW, CFS
* **`7`** = **REL SCH** = SCHEDULED RELEASE, CFS (manually reported)
* **`110** = **DIVERSN** = FLOW, CANAL DIVERSION, CFS
* **`210`** = **AUXFLOW** = FLOW AUX, CFS
* **`8`** = **FNF** = FULL NATURAL FLOW, CFS (modeled)

#### Durations (timesteps)

A particular station sensor will have data available at one or more of these duration types. The station lookup page will describe which sensor-duration combinations are available. 

* **`E`** = **EVENT** (typically 15-minute intervals)
* **`H`** = **HOURLY**
* **`D`** = **DAILY**

For example, [BCK](https://cdec.water.ca.gov/dynamicapp/staMeta?station_id=BCK) has flow data (sensor `20`) available at the 15-minute (duration `E`) and hourly (duration `H`) timesteps, as well as in a summarized daily mean format (sensor `41`, duration `D`).

### Data Access Options

#### CSV web service

`https://cdec.water.ca.gov/dynamicapp/req/CSVDataServlet?Stations={XXX}&SensorNums={XX}&dur_code={X}&Start={YYYY-MM-DD}&End={YYYY-MM-DD}`

#### Web API

Documented in PDFs linked at https://cdec.water.ca.gov/queryTools.html

#### Visual Interface

Real-time, daily, monthly query options: https://cdec.water.ca.gov/queryTools.html / https://cdec.water.ca.gov/riv_flows.html

Query most recent data by sensor for entire river basin: https://cdec.water.ca.gov/dynamicapp/getAll
https://cdec.water.ca.gov/dynamicapp/getAll?sens_num={XX}

#### R package `CDECRetrieve`

https://github.com/FlowWest/CDECRetrieve

## Spatial & Temporal Coverage

Plot and Chart that show coverage over watersheds, map here to show temporal coverage for a site # years or something 
Hilighting major limitations, full time periods missing across many watersheds 

```{r import-sensors, message=FALSE, warning=FALSE}
river_basins <- read_csv("cdec_river_basins.csv")

cdec_station_sensor_list_filename <- "cdec_station_sensor_list.Rds"

if(!file.exists(cdec_station_sensor_list_filename)){
  
  cdec_station_sensor_list <- 
    river_basins |>
      filter(selected) |>
      mutate(stations = map(river_basin, possibly(function(x) {
        CDECRetrieve::cdec_stations(river_basin = x) |> 
          select(-river_basin, -elevation) # dropping elev, CDECRetrieve issue
        }, otherwise = NA))) |> 
      unnest(stations) |>
      mutate(datasets = map(station_id, possibly(function(x) {
        CDECRetrieve::cdec_datasets(station = x)
        }, otherwise = NA))) |>
      unnest(datasets)
  
  cdec_station_sensor_list |> saveRDS(cdec_station_sensor_list_filename)
  
} else {
  
  cdec_station_sensor_list <- readRDS(cdec_station_sensor_list_filename)
  
}
```

```{r}
selected_sensors <- c(20, 41, 110)

cdec_station_sensors <- 
  cdec_station_sensor_list |> 
  mutate(min_wy = year(start %m+% months(3)),
         max_wy = year(end %m+% months(3))) |>
  filter(sensor_number %in% selected_sensors) |>
  st_as_sf(coords = c("longitude", "latitude"), crs="EPSG:4269") |>
  st_transform("EPSG:3310")
```

### Temporal Coverage

```{r plot-wy-avail, message=FALSE, warning=FALSE}
check_for_overlap <- function(wy, rb) {
  cdec_station_sensors |>
    filter(river_basin == rb) |>
    mutate(in_range = (wy >= min_wy) & (wy <= max_wy)) |>
    pull(in_range) |>
    any()
}

data_avail_by_water_year <- 
  expand_grid(water_year = seq(1949,2024,1), 
            river_basin = unique(cdec_station_sensor_list$river_basin)) |>
  mutate(has_data = map2_lgl(water_year, river_basin, check_for_overlap))

data_avail_by_water_year |>
  ggplot() + 
  geom_tile(aes(x = factor(water_year), y = factor(river_basin), fill = has_data)) +
  xlab("Water Year") + 
  ylab("CDEC River Basin") + 
  theme_minimal() + 
  theme(legend.position = "top", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  scale_y_discrete(limits=rev, position="right") +
  scale_fill_manual(values = c("FALSE" = "white", "TRUE" = "darkgray"))
```

### Spatial Coverage

```{r import-watersheds, echo=TRUE, message=FALSE, warning=FALSE, include=FALSE}
watersheds_huc8 <- 
  st_read("shp/calw221_huc_8_selected.shp", as_tibble=TRUE) |>
  janitor::clean_names() |>
  st_transform("EPSG:3310") |>
  select(huc_8, huc_8_name, wshed_id)
subwatersheds_huc12 <- 
  st_read("shp/wbd_subwatershed_selected.shp", as_tibble=TRUE) |>
  janitor::clean_names() |>
  st_transform("EPSG:3310")
subwatersheds_calw221 <-
  st_read("shp/calw221_selected.shp", as_tibble=TRUE) |>
  janitor::clean_names() |>
  st_transform("EPSG:3310")
```

```{r map-sensors}
if (interactive()) {
  leaflet::leaflet() |> 
  leaflet::addTiles() |> 
  leaflet::addPolygons(data=st_transform(watersheds_huc8, "EPSG:4326")) |>
  leaflet::addMarkers(data=st_transform(cdec_station_sensors, "EPSG:4326"))
} else {
 ggplot() + geom_sf(data=watersheds_huc8) + geom_sf(data=cdec_station_sensors)
}
```

Using watersheds from https://doi.org/10.15447/sfews.2006v4iss1art3
 
## Quality Checks 

[CONDITIONALLY NEEDED - if gage / if not already described in model section]
What quality assurance checks are implemented by monitoring agency 
What quality control checks are implemented by monitoring agency 

Error codes - May differ by station and sensor

## Data use and limitations 
(build out once we play around with the data a bit more) 
Pros and cons table for each use case 

## Questions for Data Experts 
- Please list any questions about the data source. 

