---
title: "USGS Overview"
date: "[Date]"
output:
  html_document:
     code_folding: hide
     theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(janitor)
```

## [Data Type] Summary 
The USGS is responsible for monitoring and providing information on the quantity, quality, distribution, movement, and other characteristics of surface water and groundwater resources across the United States. The NWIS Mapper provides an interface for users to explore and retrieve water data through interactive maps.

* **Source:** National Water Information System (NWIS) Mapper is a tool provided by the United States Geological Survey (USGS)
* **Accessibility:** Public, open and accessible online
* **Coverage:**  Approximately 1.9 million sites in all 50 States, the District of Columbia, Puerto Rico, the Virgin Islands, Guam, American Samoa and the Commonwealth of the Northern Mariana Islands.
Approximately 1,000 gages are operating and reporting data publicly in California. Of
these, about 60 percent are operated by the United States Geological Survey (USGS),
with the remaining gages operated by State or other agencies. A substantial number of
gages are operated by third-party entities that are not publicly reporting data on
statewide databases or lack sufficient data quality to be reported.
* **Temporal Coverage:** temporal coverage is from 1862 to present, however it varies across the different gages. Some of them have big data gaps during those years, or no data at all that goes that far back. Overall not good coverage across the system or across dates
* **Spatial Coverage:** There is at least one gage with flow data for each of the streams of interest 
* **Maintenance:** The frequency of data collection can vary depending on the specific site and the importance of the data for water resource management, flood monitoring, and other purposes.

The frequency of data collection at a particular stream gage can range from hourly to daily, and in some cases, data may be collected more frequently during specific events, such as storms or floods. Real-time data from many stream gages are available on the USGS NWIS website.
The USGS continually updates and maintains its network of monitoring sites, and data is made available to the public through the NWIS platform.
* **Contact: ** Data / Model Contact (think about if w
orth expanding to a larger list of people / orgs) - think about having table section below with projects/contacts 
* **Utilized By:** data is used for various projects, including SR JPE. 


## Spatial & Temporal Coverage
```{r}
usgs_overview <- read_csv(here::here("data-raw", "USGS_station_lookup.csv")) |> 
  clean_names()|>
  select(-percent_na_only_something_we_want_to_do_if_we_can_functionalize_this_process) |> 
  mutate(min_date = as.Date(min_date, format = "%m/%d/%Y"),
         max_date = as.Date(max_date, format = "%m/%d/%Y")) |> 
  glimpse()
```

Plot and Chart that show coverage over watersheds, map here to show temporal coverage for a site # years or something 
Hilighting major limitations, full time periods missing across many watersheds 
```{r}
ggplot(usgs_overview, aes(x = min_date, xend = max_date, y = tributary, yend = tributary)) +
  geom_segment(size = 3, color = "gray") +
  labs(title = "Date Coverage for Tributaries",
       x = "Date",
       y = "Tributary Name")

```


## Quality Checks 

[CONDITIONALLY NEEDED - if gage / if not already described in model section]
What quality assurance checks are implemented by monitoring agency 
What quality control checks are implemented by monitoring agency 

## Data use and limitations 
This data can be used and is mostly reliable, the limitations are data gaps across the system and not consistent dates/years of collection. Of the approximate 4,500 local sub-watersheds
(Hydrological Unit Code 12 [HUC12]) in California, more than 3,200 (over 70 percent) of
watersheds do not have any history of stream gaging. Less than half of the historically
gaged watersheds currently have active, publicly accessible gages today

## Questions for Data Experts 
- Please list any questions about the data source. 

